---
title: "Benchmark  of scribe"
author: IDR
date: Feb 09, 2023
output:
  md_document:
    variant: markdown_github
---

# Benchmark  of scribe

#### Disclaimer

"scribe" is a very complex program that performs many disk read-write operations as well as queries to web services.

"scribe" invokes "R CMD` check where execution times can vary greatly.

The results obtained in this test may not match your results, because they depend on parameters of the server on which it was run, speed of your internet connection and the choice of R packages.



```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)


test1 <- list(npkg = 428, 
data = data.frame(
threads=c(1,2,10,30,60,70), 
time=c(93,72,60,50,46,46)),
desc.x='Number of Threads')

test2.hpc <- list(npkg = 1086,
data = data.frame(
threads=c(30,70,300), 
time=c(84,80,46)),
desc.x='Number of Threads')

test2.local <- list(npkg = 1086,
data = data.frame(
threads=c(20), 
time=c(100)),
desc.x='Number of Threads')

test2.renv <- list(npkg = 1086,
data = data.frame(
threads=c(1), 
time=c(140)),
desc.x='Number of Threads')

test2 <- list(npkg = 1086,
data = data.frame(
threads=c("renv(1)", "scribe(030)", "scribe(070)", "scribe(300)"), 
time=c(140, 84, 80, 46)),
desc.x='Installer and number of threads')




test3.check <- list(npkg = 28,
data = data.frame(
threads=c(1, 5, 10, 20, 30), 
time=c(53, 15, 10, 7, 7)),
desc.x='Number of Threads'
)



print.graph <- function(testdata) {
  if(class(testdata$data$threads) =="numeric" ) {
    testdata$data$threads <- factor(testdata$data$threads)
  }
  
  p<-ggplot(data=testdata$data, aes(x=threads, y=time, fill=threads)) +
    labs(title = paste0('Performance on ', testdata$npkg, " packages"), x = testdata$desc.x, y = 'Time in min.')+
    geom_bar(stat="identity", fill="steelblue")+
    geom_text(aes(label=time), vjust=-0.3, size=3.5)+
    theme_minimal()
  p
}

```

#### Explanation of terms

"Number of Threads" is the maximum number of simultaneously running installations of R packages.

"Highly Dependent Package" is an R package which has a lot dependencies on other R packages.

"Non-Dependent Package" is an R package without dependencies on other R packages. 

"Low Dependent Package" is a R package which has few (1 or 2 ) dependencies on Non-dependent Packages


## Installing packages with different numbers of threads

The more threads, the shorter the installation of all packages.
This relationship is not linear.

We get to the point where the extra threads don't speed up the overall process. 
It is so, because Highly Dependent Package which has tightly coupled packages do not allow parallelisation.
The last packages in the installation list are run sequentially.

At the beginning of the installation process,  Non-dependent Package are being installed as a first packages. 
They do not have dependencies, so parallelisation is done quite effectively.


```{r, echo=FALSE}
print.graph(test1) 
```




## Installing packages and Strongly Connected Packages 

For a large number of packages, Strongly Connected Packages may appear.
Strongly Connected Packages are a group of packages, where if he wants to use one package, he will have to install a whole horde of other packages
E.g. would be "dplyr" with "tibble", "rlang", "vctrs" or "devtools" with "cli", "pkgdown", "rcmdcheck", "remotes", "roxygen2".

If there is a Highly Dependent Package then there should be Strongly Connected Packages. 

The topological sorting algorithm considers Strongly Connected Packages to parallelize effectively. 
This is not always the case, as some packages do not have a complete description of all the packages they would need for installation.
In launches, we can observe such an occurrence, when there is an increased number of threads and the installation execution time decreases significantly.


```{r, echo=FALSE}
print.graph(test2.hpc) 
```

## renv and scribe

In summary, "scribe" can install packages much faster than "renv" will ever do.
Even if, "renv" will use "pak" package for downloading packages, "scribe" will be faster, because it used parallelization for download, build, installation process.   

```{r, echo=FALSE}
print.graph(test2) 
```


## Checks in "scribe"

We missed one feature in the renv. It was running "R cmd checks" on several packages. 
We have parallelized this procedure.

Here are the result for running only checks.

```{r, echo=FALSE}
print.graph(test3.check) 
```



