---
title: "Benchmark  of scribe"
author: IDR
date: Feb 09, 2023
output:
  md_document:
    variant: markdown_github
---

# Benchmark  of scribe


## Explanation of terms

"Number of Threads" is the maximum number of simultaneously running installations of R packages.

"Highly Dependent Package" is an R package which has a lot dependencies on other R packages.

"Non-Dependent Package" is an R package without dependencies on other R packages. 

"Low Dependent Package" is an R package which has a few (1 or 2) dependencies on Non-Dependent Packages.

### Disclaimer

"scribe" is a very complex program that performs many disk read-write operations as well as queries to web services.

"scribe" invokes "R CMD` check where execution times can vary greatly.

The results obtained in this test may not match your results, because they depend on parameters of the server on which it was run, speed of your internet connection and the choice of R packages.



```{r, echo=FALSE}
library(ggplot2)
library(gridExtra)


test1 <- list(npkg = 428, 
data = data.frame(
threads=c(1,2,10,30,60,70), 
time=c(93,72,60,50,46,46)),
desc.x='Number of Threads')

test2.hpc <- list(npkg = 1086,
data = data.frame(
threads=c(30,70,300), 
time=c(84,80,46)),
desc.x='Number of Threads')

test2.local <- list(npkg = 1086,
data = data.frame(
threads=c(20), 
time=c(100)),
desc.x='Number of Threads')

test2.renv <- list(npkg = 1086,
data = data.frame(
threads=c(1), 
time=c(140)),
desc.x='Number of Threads')

test2 <- list(npkg = 1086,
data = data.frame(
threads=c("renv(1)", "scribe(030)", "scribe(070)", "scribe(300)"), 
time=c(140, 84, 80, 46)),
desc.x='Installer and number of threads')




test3.check <- list(npkg = 28,
data = data.frame(
threads=c(1,2, 3, 4, 5, 10, 15, 20, 30), 
time=c(54, 30, 22, 17, 15, 10, 8, 7, 7)),
desc.x='Number of Threads'
)



print.graph <- function(testdata) {
  if(class(testdata$data$threads) =="numeric" ) {
    testdata$data$threads <- factor(testdata$data$threads)
  }
  
  p<-ggplot(data=testdata$data, aes(x=threads, y=time, fill=threads)) +
    labs(title = paste0('Performance on ', testdata$npkg, " packages"), x = testdata$desc.x, y = 'Time in min.')+
    geom_bar(stat="identity", fill="steelblue")+
    geom_text(aes(label=time), vjust=-0.3, size=3.5)+
    theme_minimal()
  p
}

```



## Installing packages with different numbers of threads

The more threads, the shorter the time for installation of all packages.
This relationship is not linear.

We get to the point where the extra threads don't speed up the overall process. 
It is because Highly Dependent Packages (with tightly coupled packages) do not allow parallelisation.
The last packages on the installation list are run sequentially.

At the beginning of the installation process, Non-Dependent Packages are being installed first. 
They do not have dependencies, so parallelisation is done quite effectively.


```{r, echo=FALSE, fig.alt = "Performance on average number of packages"}
print.graph(test1) 
```




## Installing packages and Strongly Connected Packages 

For a large number of packages, Strongly Connected Packages may appear.
Strongly Connected Packages are a group of packages, where if one wants to use a single package, the imperative is to install a cohort of other packages.
Examples: "dplyr" with "tibble", "rlang", "vctrs" or "devtools" with "cli", "pkgdown", "rcmdcheck", "remotes", "roxygen2".

If there is a Highly Dependent Package then there should be Strongly Connected Packages. 

The topological sorting algorithm considers Strongly Connected Packages to parallelize effectively. 
This is not always the case, as some packages do not have a complete description of all the packages they require for installation.
In launches, we can observe it, when there is an increased number of threads and the installation execution time decreases significantly.


```{r, echo=FALSE, fig.alt = "Performance on large number of packages"}
print.graph(test2.hpc) 
```

## renv and scribe

In summary, "scribe" can install packages much faster than "renv" will ever do.
Even if, "renv" will use "pak" package for downloading packages, "scribe" will do it faster, because it uses parallelization for download, build, and installation process.   

```{r, echo=FALSE, fig.alt = "renv and scribe"}
print.graph(test2) 
```


## Checks in scribe

We missed one feature in the renv. It was running "R CMD checks" on several packages. 
We have parallelized this procedure.

This is the result for running only checks.

```{r, echo=FALSE, fig.alt = "Checks in scribe"}
print.graph(test3.check) 
```



